---
title: "Data_Analysis_Leaps_Subset_Selection.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Methodology from: https://rstudio-pubs-static.s3.amazonaws.com/2897_9220b21cfc0c43a396ff9abf122bb351.html
Applied to the California Housing Data Set


```{r}
# Used Google Colab to get quicker turn around on this, so had to install these packages
# Using Google Colab for R is rad: https://stackoverflow.com/questions/54595285/how-to-use-r-with-google-colaboratory
if(!require(leaps)) install.packages("leaps", repos = "http://cran.us.r-project.org")
if(!require(caret)) install.packages("caret", repos = "http://cran.us.r-project.org")
if(!require(car)) install.packages("car", repos = "http://cran.us.r-project.org")
```


```{r}
library(readr)
library(knitr)
library(caret)
library(leaps)
library(car)

housing_data = read_csv("housing.csv")
is.factor(housing_data$ocean_proximity)
housing_data$ocean_proximity = as.factor(housing_data$ocean_proximity)
is.factor(housing_data$ocean_proximity)

regsubsets.out <-
    regsubsets(median_house_value ~ .,
               data = housing_data,
               nbest = 1,       # 1 best model for each number of predictors
               nvmax = NULL,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "exhaustive")
regsubsets.out

summary.out <- summary(regsubsets.out)
as.data.frame(summary.out$outmat)
## Adjusted R2
plot(regsubsets.out, scale = "adjr2", main = "Adjusted R^2")

library(car)
layout(matrix(1:2, ncol = 2))
## Adjusted R2
res.legend <-
    subsets(regsubsets.out, statistic="adjr2", legend = FALSE, min.size = 5, main = "Adjusted R^2")
## Mallow Cp
res.legend <-
    subsets(regsubsets.out, statistic="cp", legend = FALSE, min.size = 5, main = "Mallow Cp")
abline(a = 1, b = 1, lty = 2)
res.legend

which.max(summary.out$adjr2)

summary.out$which[which.max(summary.out$adjr2),]

best.model <- lm(median_house_value ~ longitude + latitude + housing_median_age + total_rooms + total_bedrooms + population + households + median_income + I(ocean_proximity == "NEAR OCEAN") + I(ocean_proximity == "ISLAND"), data = housing_data)
summary(best.model)

```

Let's try that again with the a model with all two-way interactions.

NOTE: at least a couple of the linear combinations must be broken up or this spins for hours.

On Google Colaboratory, I noticed that there was an error message “4  linear dependencies found”

After removing two parameters `total_bedrooms` (which is colinear with `total_rooms`)  and `households` (which is colinear with `population`) this is down to “2  linear dependencies found” Reordering variables and trying again

```{r}
regsubsets2.out <-
    regsubsets(median_house_value ~ (longitude + latitude + housing_median_age + total_rooms + population + median_income + ocean_proximity) ^2,
               data = housing_data,
               nbest = 1,       # 1 best model for each number of predictors
               nvmax = 10,    # NULL for no limit on number of variables
               force.in = NULL, force.out = NULL,
               method = "exhaustive", really.big=T)
regsubsets2.out

summary2.out <- summary(regsubsets2.out)
as.data.frame(summary2.out$outmat)
## Adjusted R2
plot(regsubsets2.out, scale = "adjr2", main = "Adjusted R^2")

layout(matrix(1:2, ncol = 2))
## Adjusted R2
res.legend <-
    subsets(regsubsets2.out, statistic="adjr2", legend = FALSE, min.size = 5, main = "Adjusted R^2")
## Mallow Cp
res.legend <-
    subsets(regsubsets2.out, statistic="cp", legend = FALSE, min.size = 5, main = "Mallow Cp")
abline(a = 1, b = 1, lty = 2)
res.legend

which.max(summary2.out$adjr2)

summary2.out$which[which.max(summary2.out$adjr2),]
```


```{r}
best_model <- lm(median_house_value ~ longitude + latitude + I(ocean_proximity == "INLAND") + I(ocean_proximity == "NEAR BAY") + longitude:latitude + longitude:median_income + longitude:I(ocean_proximity == "NEAR BAY") + latitude:I(ocean_proximity == "NEAR BAY") + housing_median_age:total_rooms + housing_median_age:population + population:median_income, data = housing_data)
summary(best_model)
```

This is an interesting set of variables and makes a lot of sense. Housing prices are likely tied to the big cites (San Francisco, Los Angeles, etc.). A couple of ways to denote these are latitude, longitude and whether the location is near a Bay. Median Income may also be a proxy for the larger cities versus the more rural areas.
